---
title: "Calculate Median Fluorescence"
author: "Makayla"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
```

# 1. Read in Median Bin data for all 3 ligand conditions (NL, Fum, and Asp)
```{r}
medianBin_NL = read.csv("medianBin_NL", header = T)
medianBin_Fum = read.csv("medianBin_Fum", header = T)
medianBin_Asp = read.csv("medianBin_Asp", header = T)
```

# 2. Filter out barcodes here that don't have at least 25 total reads across all bins (and rename median, upper, and lower index to flag the ligand (ex. NL_median))
```{r}
filtered_NL = medianBin_NL %>% 
  filter(total_barcode_reads >= 25) %>% 
  rename(NL_total_bc_reads = total_barcode_reads,
         NL_lower_index = lower_index,
         NL_upper_index = upper_index,
         NL_median = median) %>% 
  select(-cumulative_p, -X) %>% #remove un-needed columns
  select(-AA_sequence) #will be added back later...

filtered_Fum = medianBin_Fum %>% 
  filter(total_barcode_reads >= 25) %>% 
  rename(Fum_total_bc_reads = total_barcode_reads,
         Fum_lower_index = lower_index,
         Fum_upper_index = upper_index,
         Fum_median = median) %>% 
  select(-cumulative_p, -X) %>%  #remove un-needed columns
  select(-AA_sequence) #will be added back later...

filtered_Asp = medianBin_Asp %>% 
  filter(total_barcode_reads >= 25) %>% 
  rename(Asp_total_bc_reads = total_barcode_reads,
         Asp_lower_index = lower_index,
         Asp_upper_index = upper_index,
         Asp_median = median) %>% 
  select(-cumulative_p, -X) %>%  #remove un-needed columns
  select(-AA_sequence) #will be added back later...
```


# 3. Merge All Ligand Conditions
## 3.1 Load consensus output data 
```{r}
#read in the file
consensus_barcodes = read.delim("extract.consensus.tsv", header = TRUE) %>% dplyr::select(-consensus_sequence_nt, - relative_freq, -cluster_size) %>% rename(AA_sequence = consensus_sequence_aa)
#nrow(consensus_barcodes) #number of barcodes
head(consensus_barcodes) #- 2 columns (barcode and Amino Acid Sequence)
```

## 3.2 merge data and re-add AA sequences and keep relevant columns
```{r}
merged = inner_join(filtered_NL, filtered_Asp, by = "barcode") %>% 
  inner_join(filtered_Fum, by = "barcode") %>%  
  inner_join(consensus_barcodes, by = "barcode") %>% 
  select(barcode, AA_sequence, NL_lower_index, NL_upper_index, NL_median, Fum_lower_index, Fum_upper_index, Fum_median, Asp_lower_index, Asp_upper_index, Asp_median) %>% 
  mutate(NL_lower_index = ifelse(is.infinite(NL_lower_index), 1, NL_lower_index), #update all lower indexes with -inf values
         Fum_lower_index = ifelse(is.infinite(Fum_lower_index), 1, Fum_lower_index),
         Asp_lower_index = ifelse(is.infinite(Asp_lower_index), 1, Asp_lower_index),)
# nrow(merged) #~56,000 barcodes
# nrow(consensus_barcodes) #~800,000 barcodes
```

# 4. Convert to MEFL
## 4.1 Load in the MEFL bin data
```{r}
# load the MEFL bin data - using the new tsv sent from Luca!
bin_to_MEFL_FA = read.delim("new_bins_to_MEFL_Fumarate-Aspartate.tsv", header = TRUE) %>% 
  rename(MEFL = X488.1..530_30.A,
         bin = X)

bin_to_MEFL_NL = read.delim("new_bins_to_MEFL_NoLigand.tsv", header = TRUE) %>% 
  rename(MEFL = X488.1..530_30.A,
         bin = X)

# Create an interpolation function to convert nondiscreet bin values to MEFL
bin_to_MEFL_inter_NL <- approxfun(bin_to_MEFL_NL$bin, bin_to_MEFL_NL$MEFL, rule = 2)
bin_to_MEFL_inter_FA <- approxfun(bin_to_MEFL_FA$bin, bin_to_MEFL_FA$MEFL, rule = 2)


# Convert from bin to MEFL
Bin_to_MEFL_all <- merged %>%
  mutate(
    NL_LowerIndex_MEFL = bin_to_MEFL_inter_NL(NL_lower_index),
    NL_UpperIndex_MEFL = bin_to_MEFL_inter_NL(NL_upper_index),
    NL_Median_MEFL = bin_to_MEFL_inter_NL(NL_median),
    Fum_LowerIndex_MEFL = bin_to_MEFL_inter_FA(Fum_lower_index),
    Fum_UpperIndex_MEFL = bin_to_MEFL_inter_FA(Fum_upper_index),
    Fum_Median_MEFL = bin_to_MEFL_inter_FA(Fum_median),
    Asp_LowerIndex_MEFL = bin_to_MEFL_inter_FA(Asp_lower_index),
    Asp_UpperIndex_MEFL = bin_to_MEFL_inter_FA(Asp_upper_index),
    Asp_Median_MEFL = bin_to_MEFL_inter_FA(Asp_median)
  )

MEFL = Bin_to_MEFL_all %>% select(barcode, AA_sequence, NL_Median_MEFL, Fum_Median_MEFL, Asp_Median_MEFL)
```


# 5. Filter WT
```{r}
# update the amino acid sequence below with the amino acid sequence of the WT
WT_MEFL = MEFL %>% filter(AA_sequence == "RHSLPYRMLRKRPMKLSTTVILMVSAVLFSVLLVVHLIYFSQISDMTRDGLANKALAVARTLADSPEIRQGLQKKPQESGIQAIAEAVRKRNDLLFIVVTDMQSLRYSHPEAQRIGQPFKGDDILKALNGEENVAINRGFLAQALRVFTPIYDENHKQIGVVAIGLELSRVTQQINDSRWSLQMAAGVKQLA")
nrow(WT_MEFL) # 1890 barcodes associated with WT AA sequence
head(WT_MEFL)
```


# 6. Get Medians for the median MEFL for WT barcodes in all 3 conditions
```{r}
# Median values for each condition for the WT
NL_Median_WT = median(WT_MEFL$NL_Median_MEFL)
Fum_Median_WT = median(WT_MEFL$Fum_Median_MEFL)
Asp_Median_WT = median(WT_MEFL$Asp_Median_MEFL)
```


# 7. Normalize by WT fluorescence and then by no ligand
```{r}
normalized_MEFL = MEFL %>% 
  mutate(Fum_wo_NL = Fum_Median_MEFL / NL_Median_MEFL,# 0. normalize by just NL
         Asp_wo_NL = Asp_Median_MEFL / NL_Median_MEFL,
         NL_wo_WT = NL_Median_MEFL/NL_Median_WT,         # 1. normalize by the WT over each of the 3 conditions
         Fum_wo_WT = Fum_Median_MEFL / Fum_Median_WT,
         Asp_wo_WT = Asp_Median_MEFL / Asp_Median_WT,
         Fum_wo_WTandNL = Fum_wo_WT - NL_wo_WT,           # 2. normalize both ligand conditions by no ligand condition
         Asp_wo_WTandNL = Asp_wo_WT - NL_wo_WT)
head(normalized_MEFL)
```

# 8. Write out to a file
```{r}
write.csv(normalized_MEFL, "./normalized_MEFL", row.names = FALSE)
```